\section{Changelog}
\begin{itemize}
\item \textbf{v.1.0} (2013/11/26) -- Initial public release.
\item v.1.1 (2013/11/26) -- Small amendment to the preamble.
\item v.1.2 (2013/11/26) -- Small amendment to the preamble including a link to the KQ category where some of the changes were taken from.
\item v.1.3 (2013/11/26) -- Added omitted attribution on Wales I\~{n}upiaq changes; alteration to changelog layout and amendment to Contact Information.
\item v.1.4 (2013/11/26) -- Forgot to update the version information in the title.
\item v.1.5 (2013/12/01) -- Added changes from Proto-Slavic to Polish, which I forgot to include in the original release. Also, added a Proto-Salish phonology I also forgot to include in the original release.
\item v.1.6 (2013/12/03) -- Added the California Vowel Shift.
\item v.1.7 (2013/12/03) -- Fixed the California Vowel Shift.
\item v.1.8 (2013/12/06) -- Credited Macska for the Pre-Slavic Vowel Changes.
\item \textbf{v.2.0} (2014/03/11) -- Added Yokuts and Lakes Plain correspondences.
\item v.2.1 (2014/03/11) -- Forgot to change version number on the first page.
\item v.2.2 (2014/03/11) -- Moved Yokuts to the Penutian group; fixed some errors.
\item \textbf{v.3.0} (2014/03/27) -- Added Northwest Caucasian, fixed Yokuts nesting errors.
\item v.3.1 (2014/03/27) -- Notes per Tropylium's request and some formatting cleanup/explanation in the section on Northwest Caucasian.
\item v.3.2 (2014/04/03) -- Added considerable information on developments in West Germanic.
\item v.3.3 (2014/04/03) -- Cross-listed some of the West Germanic developments under the Vowel Shifts section; minor fixes.
\item \textbf{v.4.0} (2014/04/03) -- Replicated the ``Most wanted sound changes" article from KneeQuickie.
\item v.4.1 (2014/04/03) -- Minor fixes.
\item v.4.2 (2014/04/04) -- Minor correction.
\item \textbf{v.5.0} (2014/04/28) -- Added some Macro-Pama-Nyungan correspondences.
\item v.5.1 (2014/04/30) -- Correction of formatting errors; change of all remaining instances of ``Linguifex'' and ``Rorschach" to ``Pogostick Man"; amendment to contact info; other minor changes; addition of Proto-Tupar\'{i} consonantal reconstruction.
\item v.5.2 (2014/05/13) -- Added Qiangic, Proto-Lolo-Burmese final -\ipa{i}(C) and -\ipa{u}(C) sequences, Paiwan, Rukai, and changes to Old Proven\c{c}al.
\item v.5.3 (2014/05/13) -- Added changes to Rhaeto-Romance.
\item v.5.4 (2014/05/28) -- Added Austronesian, Turkic, and Italic.
\item {\bf v.6.0} (2014/06/21) -- Added Northeast Caucasian and Vasconic; clarified the situation with regards to sources without a known author (these are mostly from KneeQuickie); explained policy with regards to URLs when dealing with Wikipedia pages; corrected the description of the shorthand symbol $\langle$\%$\rangle$; added a note from Tropylium on the Finno-Ugric changes.
\item v.6.1 (2014/07/18) -- Added some Austronesian changes; corrected Avestan according to comments from Alex Fink on CONLANG-L.
\item \textbf{v.7.0} (2015/01/31) -- Added some changes to the Austronesian section (including Proto-Ongan); added some changes regarding the Bantu languages and related groups; added changes for Standard German and Gothic; added correspondences for Monde languages; added Trans-New Guinea section. Cleaned up the Key to Abbreviations.
\item v.7.1 (2015/01/31) -- Corrected an accidentally omitted attribution.
\item v.7.2 (2015/01/31) -- Corrected a typo in the changelog.
\item v.7.3 (2015/01/31) -- Corrected an oversight in formatting.
\item v.7.4 (2015/02/01) -- Formatting and attribution fixes. Added reconstructed Proto-Trans New Guinea phonology.
\item v.7.5 (2015/02/03) -- Correcting omissions, including an acknowledgment that a quote from Whimemsz is sourced from KneeQuickie.
\item v.7.6 (2015/02/25) -- Fixed misspelling of ``Falsican"; fixed an alphabetical ordering 
error in the Indo-European section.
\item v.7.7 (2015/06/02) -- Added missing section on the development of Cheyenne that I forgot to add to the document originally; added a lot of potential Bantu correspondences (see the note in the introduction to the Niger-Congo section regarding said correspondences); added Faroese vowel shift information; added correspondences for some Kalamian languages; added changes from Proto-Oceanic to Hiw.
\item {\bf v.8.0} (2015/11/16) -- Added Piscataway, Mi'kmaq, and Cree correspondences to the Algonquian section; added Kainantu-Goroka correspondences to the Trans-New Guinea section; added Vandalic correspondences; added some rGyalrongic and Tibetic correspondences to the Sino-Tibetan section; added Philippine and Oceanic correspondences to the Austronesian section; added Pai correspondences to the Yuman-Cochim\'{\i} section; added Chumashan correspondences; added Ofai\'{e}-J\^{e} correspondences; added Bakairi correspondences; added Lenmichian correspondences; added some Na-Dene correspondences and moved the Athabaskan section under Na-Dene. Fixed a formatting error in the changelog.
\item v.8.1 (2015/11/16) -- Corrected some formatting errors.
\item v.8.2 (2016/02/18) -- Added missing section on Cheyenne that was lost during the move from v.7.7 to v.8.1. Corrected some errors in the changeling.
\end{itemize}
\clearpage

